Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.Rmd
update meeting note
change the first plot to a faceted plot.
Remove axis and gridline from many plots.

```{r get-lineup-data}
# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))
  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
}
```

@@

@@

patchwork::wrap_plots(

  # Plot base data
  mod$plot(dat, theme = theme_light()) +
    ggtitle("A. Good residuals") + 
    xlab("Fitted values") + 
    ylab("Residuals"),

  # Include the z term
  # Reuse the x and e terms
  mod$plot(mod$
             set_prm("include_z", TRUE)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("B. Non-linearity") +
    xlab("Fitted values") + 
    ylab("Residuals"),

  # Build a heter model
  # Reuse the x and e terms
  mod$plot(heter_model(b = 64)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("C. Heteroskedasticity") +
    xlab("Fitted values") + 
    ylab("Residuals"),

  # Build a heter model with non-normal error
  # Reuse the x term
  mod$plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
             gen(300, computed = select(dat, x)), 
    theme = theme_light()) + 
    ggtitle("D. Non-normality") +
    xlab("Fitted values") + 
    ylab("Residuals"),
  ncol = 4
)
```

Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.

@@

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at Figure \ref{fig:residual-plot-common-departures}B. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis. Visually, it sometimes results in the so-called "butterfly" shape as shown in the Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution. A favourable graphical summary for this task is the quantile-quantile plot. As we mainly discuss residual plots, non-normality will not be the focus of this paper. For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D. When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated. For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.

@@

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(shape = factor(shape, levels = shape_labels)) %>%
  VI_MODEL$plot() +
  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Skewed", 
                                    "Discrete"))) %>%
  VI_MODEL$plot() +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank()) +
  facet_wrap(~x_dist, ncol = 4, scales = "free")
```

@@

  # Combined with data for 300 observations
  bind_rows(dat_n_300) %>%
  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
  VI_MODEL$plot() +
  theme_light() +
  facet_wrap(~n, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for sigma 0.5
  bind_rows(dat_sigma_05) %>%
  VI_MODEL$plot() + 
  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a, levels = a_labels)) %>%
  VI_MODEL$plot() +
  facet_wrap(~a, scales = "free", ncol = 3) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for b = 0.25
  bind_rows(dat_b025) %>%
  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
  VI_MODEL$plot() +
  facet_wrap(~b, scales = "free", ncol = 5) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

>>>>

```{r get-lineup-data}
# Create the dir folder
if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))

# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
}
```

@@

(change it to facet)

(remove backticks for some plots)

@@

bind_rows(
  dat %>%
    mutate(type = "Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "Non-normality")
) %>%
  mod$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~type, ncol = 4, scales = "free") +
  xlab("Fitted values") +
  ylab("Residuals")
```

Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. The first panel of Figure \ref{fig:residual-plot-common-departures} shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.

@@

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the second panel of Figure \ref{fig:residual-plot-common-departures}. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis. Visually, it sometimes results in the so-called "butterfly" shape as shown in the third panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution. A favourable graphical summary for this task is the quantile-quantile plot. As we mainly discuss residual plots, non-normality will not be the focus of this paper. For a consistent comparison, the residual plot of this departure is still presented in the fourth panel of Figure \ref{fig:residual-plot-common-departures}. When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated. For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in the fourth panel of Figure \ref{fig:residual-plot-common-departures}.

@@

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(shape = factor(shape, levels = shape_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Skewed", 
                                    "Discrete"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank()) +
  facet_wrap(~x_dist, ncol = 4, scales = "free")
```

@@

  # Combined with data for 300 observations
  bind_rows(dat_n_300) %>%
  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~n, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for sigma 0.5
  bind_rows(dat_sigma_05) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) + 
  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a, levels = a_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~a, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

@@

  # Combined with data for b = 0.25
  bind_rows(dat_b025) %>%
  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~b, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme(axis.text = element_blank())
```

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.tex
change the first plot to a faceted plot.
Remove axis and gridline from many plots.

Graphical summaries in which residuals are plotted against fitted values
or other functions of the regressors that are approximately orthogonal
to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. Figure
\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
with residuals evenly distributed at both sides of the horizontal zero
line, with no noticeable patterns.

@@

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values in the
residual plot could be indicative of non-linearity. An example residual
plot containing visual pattern of non-linearity is given at Figure
\ref{fig:residual-plot-common-departures}B. One can clearly observe the
``S-shape'' from the residual plot as the cubic term is not captured by
the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in the Figure \ref{fig:residual-plot-common-departures}C, or the
``left-triangle'' and ``right-triangle'' shape where the smallest
variance occurs at one side of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is
usually harder to detect from a residual plot since a scatter plot do
not readily reveal the marginal distribution. A favourable graphical
summary for this task is the quantile-quantile plot. As we mainly
discuss residual plots, non-normality will not be the focus of this
paper. For a consistent comparison, the residual plot of this departure
is still presented in Figure \ref{fig:residual-plot-common-departures}D.
When the number residuals below and above the horizontal axis are uneven
across the local regions along the \(x\)-axis, we expect that the
normality assumption is violated. For example, given a skewed error
distribution, there will be fewer data points and more outliers on one
side of the horizontal axis as shown in Figure
\ref{fig:residual-plot-common-departures}D.

>>>>

Graphical summaries in which residuals are plotted against fitted values
or other functions of the regressors that are approximately orthogonal
to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. The first panel of Figure
\ref{fig:residual-plot-common-departures} shows an ideal residual plot
with residuals evenly distributed at both sides of the horizontal zero
line, with no noticeable patterns.

@@

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values in the
residual plot could be indicative of non-linearity. An example residual
plot containing visual pattern of non-linearity is given at the second
panel of Figure \ref{fig:residual-plot-common-departures}. One can
clearly observe the ``S-shape'' from the residual plot as the cubic term
is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in the third panel of Figure
\ref{fig:residual-plot-common-departures}, or the ``left-triangle'' and
``right-triangle'' shape where the smallest variance occurs at one side
of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is
usually harder to detect from a residual plot since a scatter plot do
not readily reveal the marginal distribution. A favourable graphical
summary for this task is the quantile-quantile plot. As we mainly
discuss residual plots, non-normality will not be the focus of this
paper. For a consistent comparison, the residual plot of this departure
is still presented in the fourth panel of Figure
\ref{fig:residual-plot-common-departures}. When the number residuals
below and above the horizontal axis are uneven across the local regions
along the \(x\)-axis, we expect that the normality assumption is
violated. For example, given a skewed error distribution, there will be
fewer data points and more outliers on one side of the horizontal axis
as shown in the fourth panel of Figure
\ref{fig:residual-plot-common-departures}.
